\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-02-07 paper/main.tex
edit model sec

\subsection{Model}
Before introducing the models core components, we will clarify the previously mentioned notions of top-down and bottom-up. We will refer to features and methods, starting from the observable characteristics of either drug or protein inferring the likely molecular mechanisms as top-down. On the other hand, bottom-up refers to approaches deducing from even finer molecular properties, such as single functional groups or amino acid sequences for drugs and proteins, respectively.

In order to build a method that incorporates both top-down and
bottom-up features, we first created a model for each
individually. Hence, we assemble rich molecular structure based
features for drugs from
\textit{SmilesTransformer} \citep{SmilesTransformer} and proteins from
\textit{DeepGOPlus} \citep{DeepGoPlus}. \textit{SmilesTransformer}
introduces an autoencoder, learning over the SMILES strings and thus
the molecular organization of each drug in an unsupervised manner. On
the other hand, we take advantage of the pretrained models of
\textit{DeepGOPlus}, obtaining features from the proteins amino acid
sequence, showing significant performance in the field of protein
function prediction. Thus, both embeddings seem to suitably supplement
the following ontology based representations.

In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
obtain ontology based representations. DL2vec constructs a graph by
introducing vertices and edges for each ontology class and axiom,
respectively, followed by random walks starting from each
entity. These walks are eventually learned on using a Word2vec
\citep{Word2vec2013} model. Thus, we pick up rich, neighbourhood
focused representations for each entity, which has shown great results
for representing protein function and phenotypes. The overall
structure of the ontology can be seen in Figure~\ref{fig:Onto}.

@@

\subsubsection{Siamese neural networks and modular learnable feature transformation}
As we want to learn from the similarity of drug side effects and
protein phenotypes, we use a deep Siamese neural network.
A Siamese neural network aims to learn a similarity between two
embeddings.

>>>>

\subsection{Model}

Our model combines ``top-down'' and ``bottom-up'' information for
drug--target identification. We consider an approach to be
``top-down'' when observable characteristics of either a drug (such as
a drug effect) or protein (such as a protein function, or phenotypes
resulting from a loss of function) are used to provide information
about a molecular mechanisms; we consider an approach ``bottom-up''
when structural or other molecular information is used to determine a
mechanism.  In order to build a method that incorporates both top-down
and bottom-up features, we first create a model for each type of
feature separately.  As features for the bottom-up model, we use
features derived from molecular structures of drugs from the
\textit{SmilesTransformer} \citep{SmilesTransformer} and molecular
features for proteins from \textit{DeepGOPlus}
\citep{DeepGoPlus}. \textit{SmilesTransformer} introduces an
autoencoder, learning over the SMILES strings and therefore the
molecular organization of each drug in an unsupervised manner. 
\textit{DeepGOPlus} provides  features derived from protein amino acid
sequences which are useful to predict protein function.
% Thus, both embeddings seem to suitably supplement
% the following ontology based representations.
In the top-down section, we used \textit{DL2vec} \citep{DL2vec2020} to
obtain ontology based representations. DL2vec constructs a graph by
introducing vertices and edges for each ontology class and axiom,
respectively, followed by random walks starting from each
entity. These walks are eventually learned on using a Word2vec
\citep{Word2vec2013} model. Thus, we pick up rich, neighbourhood
focused representations for each entity, which has shown great results
for representing protein function and phenotypes. The overall
structure of the ontology can be seen in Figure~\ref{fig:Onto}.

@@

\subsubsection{Siamese neural networks and modular learnable feature transformation}

As we want to learn from the similarity of drug side effects and
protein phenotypes, we use a deep Siamese neural network.
A Siamese neural network aims to learn a similarity between two
embeddings.
